---
title: "Analysis on the relationship between the Sea Dispersal Potential index and morphological traits of Galápagos littoral plants"
output: html_notebook
Authors: Fuster-Calvo, A. Nogales, M. Heleno, R. Vera, C. Vargas, P.
---

## Script description

1. Testing the relationship between morphological traits related to thalassochory plus additional diaspore features (fleshiness, shape and size) and the SDPi.

 1.1. Mann–Whitney–Wilcoxon tests performed to analyze whether SDPi values differ between species classified as thalassochorous vs. non-thalassochorous
 1.2. Linear Models fitted to analyse the relationship between shape and size of diaspores and the SDPi values.


```{r}
library(geoR)
library(ggplot2)
library(rcompanion)
library(lmtest)
library(hrbrthemes)
library(ggplot2)
library(dplyr)
library(MKinfer)
library(MASS)
library(aods3)
library(visreg)
```


```{r}
df <- read.csv("SDP-traits.csv", header = TRUE)

colnames(df)[colnames(df) == "length_mean_mm"] <- "length"

structure(df)
```



#  (1) Comparing SDPi values between THA vs. non-THA species 

```{r}
# Analysis of variables - normality

plotNormalHistogram(df$SDP_t33.THA)
plotNormalHistogram(df$SDP_t33.noTHA)

plotNormalHistogram(df$SDP_t21.THA)
plotNormalHistogram(df$SDP_t21.noTHA)


shapiro.test(df$SDP_t33.THA) #p-value = 0.02547
shapiro.test(df$SDP_t33.noTHA) #p-value = 0.4649

shapiro.test(df$SDP_t21.THA) #p-value = 0.007653
shapiro.test(df$SDP_t21.noTHA) #p-value = 0.000131

```


The sampled substets of populations don't follow a normal distribution for SDPi values.
We use the non-parametric Mann–Whitney–Wilcoxon test: 

```{r}

summary(df$SDP_t33.THA) # Median = 71.75
summary(df$SDP_t33.noTHA) #Median = 31.89

summary(df$SDP_t21.THA) # Median = 13.35
summary(df$SDP_t21.noTHA) # Median = 0


wilcox.test(df$SDP_t33 ~ df$Syndrome_THA, data = df) # p-value = 0.09533 (U = 16.5)
wilcox.test(df$SDP_t21 ~ df$Syndrome_THA, data = df) # p-value = 0.02711 (U = 11)

#In t33h, the distributions in the two groups (THA vs no-THA) don't differ.
#In t21d, the distributions in the two groups differed significantly.



# Do species categorized as THA have higher SDP than no-THA in t21d?

wilcox.test(df$SDP_t21 ~ df$Syndrome_THA, data = df, 
        exact = FALSE, alternative = "less") # p-value = 0.01356

# #In t21d, the median of species THA is higher than species no-THA.
```


## Boxplot THA vs. no-THA

```{r}

## SDP t33

plot_SDP33_syndromes<-ggplot(df , aes(x=Syndrome_THA, y=SDP_t33),position = position_dodge(5)) + 
    geom_boxplot() +
   geom_dotplot(binaxis='y', stackdir='center', dotsize=0.8, fill="black") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)) +
    ggtitle("SDPt33~Syndrome") +
    xlab("")+
scale_fill_manual(values = c("gray87", "lightblue"))+
  stat_boxplot(geom ='errorbar', width = 0.6) +
   stat_summary(fun.y=mean, geom="point", shape=10, size=6, color="red")+
scale_y_continuous(sec.axis = dup_axis(label = NULL,
                                         name = NULL),
                     expand = expand_scale(mult = c(0, 0)),
                     breaks = pretty(c(0,100), n = 5),
                     limits = c(-3,102))



## SDP t21

plot_SDP21_syndromes<-ggplot(df , aes(x=Syndrome_THA, y=SDP_t21),position = position_dodge(5)) + 
    geom_boxplot() +
   geom_dotplot(binaxis='y', stackdir='center', dotsize=0.8, fill="black") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)) +
    ggtitle("SDPt21~Syndrome") +
    xlab("")+
scale_fill_manual(values = c("gray87", "lightblue"))+
  stat_boxplot(geom ='errorbar', width = 0.6) +
  stat_summary(fun=mean, geom="point", shape=10, size=6, color="red")+
scale_y_continuous(sec.axis = dup_axis(label = NULL,
                                         name = NULL),
                     expand = expansion(mult = c(0, 0)),
                     breaks = pretty(c(0,100), n = 5),
                     limits = c(-3,102))


par(mfrow=c(1,2))
plot(plot_SDP33_syndromes)
plot(plot_SDP21_syndromes)

```



# (2) Comparing SDPi values between species with Fleshy vs. Dry diaspores.

```{r}
# Analysis of variables - normality

plotNormalHistogram(df$SDP_t33_Fleshy)
plotNormalHistogram(df$SDP_t33_Dry)

plotNormalHistogram(df$SDP_t21_Fleshy)
plotNormalHistogram(df$SDP_t21_Dry)


shapiro.test(df$SDP_t33_Fleshy) #p-value = 0.1216
shapiro.test(df$SDP_t33_Dry) #p-value = 0.1582

shapiro.test(df$SDP_t21_Fleshy) #p-value = 0.003459
shapiro.test(df$SDP_t21_Dry) #p-value = 0.008011

# for t33h, Fleshy and Dry distribute normally.
# for t21d, they don't.

```


## Mann–Whitney–Wilcoxon tests

```{r}
#t33h
summary(df$SDP_t33_Fleshy) # Median = 70.06
summary(df$SDP_t33_Dry) # Median = 35.60
wilcox.test(df$SDP_t33 ~ df$Fleshiness, data = df) # W = 43, p-value = 0.967

# t.test (parametric) on t33h to check whether results differ from Mann-Whitney U test.
t.test(df$SDP_t33~df$Fleshiness) #t = -0.1517, df = 13.687, p-value = 0.8816
# mean in group Dry (51.94273) mean in group Fleshy (54.65875).


#t21d
summary(df$SDP_t21_Fleshy) # Median = 8.37
summary(df$SDP_t21_Dry) # Median = 8.39
wilcox.test(df$SDP_t21 ~ df$Fleshiness, data = df) # W = 46, p-value = 0.8999
```



# ############ SDP index ~ Morphological traits ############



## Variables transformations

Variables visualization
```{r}

par(mfrow=c(2,2))
###SDP index t33h
SDP33_normHist<-plotNormalHistogram(df$SDP_t33)

###SDP index t21d
SDP21_normHist<-plotNormalHistogram(df$SDP_t21)

###Length
length_normHist<-plotNormalHistogram(df$length)

###Shape
Shape_normHist<-plotNormalHistogram(df$shape_EI)

```


Transformation - Tukey’s Ladder of Powers

transformTukey() function performs iterative Shapiro–Wilk tests, and finds the lambda value that maximizes the W statistic from those tests.  In essence, this finds the power transformation that makes the data fit the normal distribution as closely as possible with this type of transformation.

```{r}
###SDP index t33h
SDP33_Tuk<-transformTukey(df$SDP_t33) #Lambda = 0.9

###SDP index t21d
SDP21_Tuk<-transformTukey(df$SDP_t21) #Lambda = 0.33

###Length
length_Tuk<-transformTukey(df$length) #Lambda = -0.125

###Shape
shape_tuk<-transformTukey(df$shape_EI) #Lambda = -2.95

```


## Linear Models. SDPi ~ length + shape
```{r}


lm_SDP33_1<-lm(data=df, SDP_t33~length_Tuk+shape_tuk, na.action=na.exclude) # meet assumptions
lm_SDP21_2<-lm(data=df, SDP21_Tuk~length_Tuk+shape_tuk, na.action=na.exclude) # meet assumptions

# ----------------- Model evaluation

### ---- Residuals plots

### SDPt33h
par(mfrow=c(2,2))
plot(lm_SDP33_1)

### SDPt21d
par(mfrow=c(2,2))
plot(lm_SDP21_2)



### ---- Assumptions' tests

#linearity
resettest(lm_SDP33_1)#p-value = 0.8486
resettest(lm_SDP21_2) #p-value = 0.9421

#normality
shapiro.test(residuals(lm_SDP33_1))#p-value = 0.7909
shapiro.test(residuals(lm_SDP21_2))#p-value = 0.3551

#Homocedasticity
lmtest::bptest(lm_SDP33_1) #p-value = p-value = 0.9752
lmtest::bptest(lm_SDP21_2)#p-value =  0.2656




summary(lm_SDP33_1)
summary(lm_SDP21_2)
```




## Data visualization. SDPi ~ length

```{r}
## SDPi t33h ~ length

plot_SDP33_length<-ggplot(df, aes(x=length_Tuk, y=SDP33_Tuk)) + 
  theme_ipsum()+
  geom_point()+
  geom_smooth(method=lm)



## SDPi t21d ~ length

plot_SDP21_length<-ggplot(df, aes(x=length_Tuk, y=SDP21_Tuk)) + 
  theme_ipsum()+
  geom_point()+
  geom_smooth(method=lm)


par(mfrow=c(1,2))
plot_SDP33_length
plot_SDP21_length

```




## #### Flotability and Viability components of SDPt33 ~ Length

Length was significantly related with SDP t33h. In the folowing lines, it is analyzed the relationship between diaspore's length and both components of the SDP t33h (floatability and viability). 

## Linear Models

```{r}
lm_SDP33_floatability<-lm(data=df, Ft33~length_Tuk, na.action=na.exclude)
lm_SDP33_viability<-lm(data=df, Vt33~length_Tuk, na.action=na.exclude)


### ---- Residuals plots

#SDPt33h
par(mfrow=c(2,2))
plot(lm_SDP33_floatability)


#SDPt21d
par(mfrow=c(2,2))
plot(lm_SDP33_viability)


### ---- Assumptions' tests

#linearity

resettest(lm_SDP33_floatability)#p-value = 0.8604
resettest(lm_SDP33_viability) #p-value = 0.7949

#normality

shapiro.test(residuals(lm_SDP33_floatability))#p-value = 0.3147
shapiro.test(residuals(lm_SDP33_viability))#p-value = 0.8269

#Homocedasticity

lmtest::bptest(lm_SDP33_floatability) #p-value = p-value = 0.2848
lmtest::bptest(lm_SDP33_viability)#p-value = 0.1845



summary(lm_SDP33_floatability)
summary(lm_SDP33_viability)
```


## Data visualization

```{r}
# Ft33 ~ log(length)

plot_Ft33_length<-ggplot(df, aes(x=length_Tuk, y=Ft33)) + 
  theme_ipsum()+
  geom_point()+
  geom_smooth(method=lm)



# Vt33 ~ log(length)

plot_Vt33_log.length<-ggplot(df, aes(x=length_Tuk, y=Vt33)) + 
  theme_ipsum()+
  geom_point()+
  geom_smooth(method=lm)+
  ylim(0,150)



par(mfrow=c(1,2))
plot_Ft33_length
plot_Vt33_log.length

```









