---
title: "Analysis on the relationship between the Sea Dispersal Potential index and distribution of littoral plants across the Galápagos "
output: html_notebook
Authors: Fuster-Calvo, A. Nogales, M. Heleno, R. Vera, C. Vargas, P.
---


## Script description

1. Generalized Linear Models are fitted between SDPi and the distribution of species across the Galápagos (N islands where present).

 1.1. GLM models considering the islands in the present.
 1.2. GLM models considering the paleo-islands configuration.


```{r}
library(MASS)
library(visreg)
library(aods3)
library(lmtest)
library(rcompanion)
library(ggplot2)
library(DHARMa)
library(AER)
library(performance)
```

```{r}
GLM <- read.csv("../Data/GLM_SDP.csv",header=TRUE)
```


## 1.1. GLM models considering the islands' configuration in the present. 

```{r}

glm <- glm(nislands ~ SDP33 + SDP21,GLM, family=poisson)



#---- Residuals' plot

par(mfrow=c(2,2))
plot(glm)


#---- Linearity in each variable

ggplot(data.frame(x1=GLM$SDP33,pearson=residuals(glm,type="pearson")),
      aes(x=x1,y=pearson)) +
    geom_point() +
    theme_bw()

ggplot(data.frame(x1=GLM$SDP21,pearson=residuals(glm,type="pearson")),
      aes(x=x1,y=pearson)) +
    geom_point() +
    theme_bw()

#---- LNormality of residuals
qqnorm(residuals(glm))





### ---- Assumptions' tests

#linearity
resettest(glm)#p-value = 0.04569

#normality
shapiro.test(residuals(glm))#p-value = 0.3004

#Homocedasticity
lmtest::bptest(glm) #p-value = 0.05929


#Dispersion
dispersiontest(glm,trafo=1) #p-value = 0.8
check_overdispersion(glm) #p-value =  0.533


anova(glm)
summary(glm)
```




## 1.2. GLM models considering the paleo-islands' configuration.


```{r}

glm_paleo<- glm(paleoislands ~ SDP33 + SDP21,GLM, family=poisson)

glm_paleo3<- glm(paleoislands ~ SDP33 + SDP21,GLM, family=quasipoisson)



### ---- Residuals plots

par(mfrow=c(2,2))
plot(glm_paleo)

#---- Residuals' plot

par(mfrow=c(2,2))
plot(glm)


#---- Linearity in each variable

ggplot(data.frame(x1=GLM$SDP33,pearson=residuals(glm_paleo,type="pearson")),
      aes(x=x1,y=pearson)) +
    geom_point() +
    theme_bw()

ggplot(data.frame(x1=GLM$SDP21,pearson=residuals(glm_paleo,type="pearson")),
      aes(x=x1,y=pearson)) +
    geom_point() +
    theme_bw()

#---- LNormality of residuals
qqnorm(residuals(glm_paleo))




### ---- Assumptions' tests

#linearity
resettest(glm_paleo)#p-value = 0.01243**

#normality
shapiro.test(residuals(glm_paleo))#p-value = 0.3105

#Homocedasticity
lmtest::bptest(glm_paleo) #p-value = 0.3282

#Dispersion
dispersiontest(glm_paleo,trafo=1) #p-value = 1
check_overdispersion(glm_paleo) #p-value = 0.959

anova(glm_paleo)
summary(glm_paleo)
```


Model "glm_paleo" violates the assumption of linearity (P = 0.01243). We therefore transform the predictor SDP21, which does not follow a normal distribution. We use a power transformation using Tukey's Ladder of Powers method. transformTukey() function performs iterative Shapiro-Wilk tests, and finds the lambda value that maximizes the W statistic from those tests.  In essence, this finds the power transformation that makes the data fit the normal distribution as closely as possible with this type of transformation.

The model is fitted again with the transformed variable SDPi21d_Tuk (SDP21d^0.325).

```{r}
SDPi21d_Tuk<-transformTukey(GLM$SDP21) # lambda = 0.325

glm_paleo1<- glm(paleoislands ~ SDP33 + SDPi21d_Tuk,GLM, family=poisson)


### ---- Residuals plots

par(mfrow=c(2,2))
plot(glm_paleo1)


#---- Linearity in each variable

ggplot(data.frame(x1=GLM$SDP33,pearson=residuals(glm_paleo1,type="pearson")),
      aes(x=x1,y=pearson)) +
    geom_point() +
    theme_bw()

ggplot(data.frame(x1=GLM$SDP21,pearson=residuals(glm_paleo1,type="pearson")),
      aes(x=x1,y=pearson)) +
    geom_point() +
    theme_bw()

#---- LNormality of residuals
qqnorm(residuals(glm_paleo1))



### ---- Assumptions' tests

#linearity
resettest(glm_paleo1)#p-value = 0.9488

#normality
shapiro.test(residuals(glm_paleo1))#p-value = 0.9196

#Homocedasticity
lmtest::bptest(glm_paleo1) #p-value = 0.5572

#Dispersion
dispersiontest(glm_paleo1,trafo=1) #p-value = 1
check_overdispersion(glm_paleo1) #p-value = 0.987

```





